### 2025-02-25 draft.tex
caption changes

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{biblatex} % Imports biblatex package
\addbibresource{references.bib}
\usepackage{changepage}
\usepackage{parskip}

\title{Creating More Meaningful Explanations for Audio Deepfake Detection}
\date{2024-10-01} 
\author{Jacob LaRock\\1667321}
\linespread{1.5}

\def\Institute{Universität Siegen}
\def\KindOfWork{Bachelorarbeit}
\def\Studiengang{Wirtschaftsinformatik}
\def\Fakultaet{Fakultät III: Wirtschaftswissenschaften, Wirtschaftsinformatik und Wirtschaftsrecht}

\def\Title{Explaining Audio Deepfake Detection with Perceptible Features and LIME}
\def\Subtitle{}

\def\student{Jaocb LaRock}
\def\studentno{1667321}
\def\place{Siegen}
\def\Date{17.03.2025} 
\def\semester{6}

\def\erstpruefer{Gunnar Stevens}
\def\erstprueferMail{Gunnar.Stevens@uni-siegen.de}

\makeatletter
\newcommand\notsotiny{\@setfontsize\notsotiny\@vipt\@viipt}
\makeatother

\begin{document}
	\begin{titlepage}
        \begin{minipage}{0.9\linewidth}
			\centering
			\includegraphics [width=0.35\textwidth]{images/LogoSiegen}
        \end{minipage}

		\vspace{2cm}

		\centering

		{\Large\bfseries \Title\par}
		{\large\bfseries \Subtitle\par}
		\vspace{2cm}

		{\large \textbf{\KindOfWork}\par}

		\vspace{1.5cm}

		{\normalsize im Studiengang \\
			\Studiengang \ im \semester. Semester \\
		}
		\vspace{0.5cm}
		{\normalsize vorgelegt von}
		\vspace{0.5cm}

		{\normalsize \textbf{\student} \\
			Matr.-Nr.: \studentno  \\}
		{am \normalsize \Date\par}

		\vspace{0.5cm}

		{\Institute} \par
		{\Fakultaet}\par

		\vspace{1cm}

    {\begin{table}[h]
			\centering
			\begin{tabular}{lll}
				{Erstprüfer:} & \erstpruefer & \erstprueferMail \\
			\end{tabular}
		\end{table}}     
		\vfill
	\end{titlepage}
    \newpage
    \section{Abstract}
	\sloppy
	With the unprecedented advancement of Generative Artificial Intelligence (GenAI), the threat
	of voice scams using synthetic voices has become a serious concern across various sectors.
	Recent efforts have focused on identifying fake voices through handcrafted features, deep
	learning models, and hybrid approaches. However, most existing methods lack explainability,
	rendering their predictions non-transparent to users. This paper proposes a novel,
	interpretable, and transparent method for fake voice identification by introducing a hybrid
	deep learning model that leverages multiple extracted features. The hybrid model consists of
	two main components: the first component addresses heterogeneous feature spaces by employing
	deep convolutional sub-models tailored to individual features, while the second component, the
	terminus model, utilizes the concatenated representations from the final layers of each
	sub-model as input. The terminus model follows a typical multi-layer perceptron architecture,
	enabling effective integration and classification of the diverse feature representations. To
	enhance interpretability, we decompose the model’s decisions using Local Interpretable
	Model-agnostic Explanations (LIME), exploiting the identical feature representation before the
	concatenation layers to address challenges related to multi-dimensional feature
	representations. To evaluate the importance and trustworthiness of features in the generated
	explanations, we propose two metrics: importance and trust. Extensive experiments are
	conducted on the In-the-Wild dataset, which is designed to test the generalization capability
	of synthetic audio detection methods. The experimental results demonstrate that our approach
	achieves performance comparable to benchmark methods. Furthermore, the results based on our
	proposed metrics conclude that certain perceptible features demonstrate promise for generating
	explanations that are meaningful to general users. For reproducibility, the source code for
	these experiments is available in the following repository:
	\url{https://github.com/jacoblarock/fake_voices_xai}
    \section{Introduction}
    As the use of generative methods for the creation of synthetic voices becomes more widespread,
    so grows the need for reliable and usable detection methods for the protection of the security
    of people and businesses alike. In particular, the rise of deepfake technology has led to
	concerns about its potential misuse in areas such as politics, entertainment and national
	security by, for instance, potentially allowing malicious actors to create fake audio
	recordings that appear to be genuine statements made by public figures or to create
	genuine-seeming recordings of events that never happened, which could have significant
	consequences if used to spread misinformation, from defaming individuals to creating political
	tension \cite{veerasamy_rising_2022,albahar_deepfakes_2005}.
	\par
	The field of Explainable AI (XAI) shows promise for producing useful and interpretable results
	from such models. Explainable AI refers to the ability of artificial intelligence systems,
	such as machine learning models and neural networks, to provide understandable and
	interpretable explanations for their decisions or predictions \cite{hind_explaining_2019}.
	This means that XAI systems can articulate why they made a particular decision, what factors
	influenced it and how confident they are in their conclusion \cite{hind_explaining_2019}. For
	the detection of audio deepfakes, this could mean that the model can justify its
	classification with feature-based evidence, allowing for verification of the result by the end
	user. Much of the existing research, further discussed in the next section, follows the path
	of producing a model with the best possible benchmarks against datasets of samples, without
	taking into account if the results can be usefully interpreted, while existing explorations
	into making audio deepfake detection explainable have produced results that often require a
	lot of background knowledge to meaningfully interpret. Aside from this, explainability in the
	area of audio deepfake detection remains an open challenge \cite{cuccovillo_open_2022}, one
	even promoted by regulatory bodies such as the European Union with the "right to explain"
	\cite{goodman_european_2017}.
	\par
	In this work, we make a distinction of two main categories of input features for a synthetic
	voice detection model: perceptible and imperceptible features. Perceptible features are
	features that can be perceived by the human ear, often vocal qualities such as jitter, shimmer
	or pitch fluctuation that have a wide range of uses even outside of audio classification such
	as diagnosis of disease \cite{chaiwongyen_deepfake-speech_2023}, while imperceptible features
	are typically out of the range of human hearing, and may not directly reflect a vocal quality,
	otherwise referred to as "speaker-independent" features \cite{liu_hidden--wave_2023}. We use a
	new combination of features for our method, with an emphasis being placed on perceptible
	features, as we believe these to be the most likely to be understood by the layperson.
	\par
	We introduce a hybrid deep-learning model, with the purpose of combining these
	multidimensional features into a single output. The hybrid model consists firstly of a
	component sub-model for each of the input features, in order to address the problem of varying
	dimensionality of the extracted features without the drawbacks of costly transformations in
	the pre-processing phase. The final part of our hybrid model is what we will refer to as the
	terminus model, a model that uses the concatenated outputs of the individual sub-models as
	inputs and distills down to a singular output with a classic multi-layer perceptron
	architecture, allowing for an effective combination of the features into a final singular
	classification result.
	\par
	In order to introduce explainability into our method, we generate explanations using a method
	that first generated the intermediate output data from the sub-models and then applies the
	Local Interpretable Model-agnostic Explanations (LIME) \cite{ribeiro_why_2016} on the terminus
	model, which allows us to, for an individual classification, assess the impact of each input
	feature on the final result through the local explanation, as, due the independence of the
	sub-models from one another, the importance of their outputs directly correlates with the
	importance of their inputs when it comes to the end-classification made by the terminus model.
	\par
	We then stood before the problem of assessing our explanations, more specifically, assessing
	the impact of the input features on the average local explanation, in order to make a
	conclusion about the usefulness or promise of our chosen features for providing understandable
	explanations of our model's classifications. In order to tackle the problem with our local
	explanations, and due to a lack of metrics to assess the aggregated average value of local
	explanations, we also introduce two metrics for such an aggregate evaluation of large number
	of generated LIME explanations, importance and trust, on a per-feature basis, which we use to
	make a conclusion about the value of the selected feature set for the purpose of sensible
	explanations.
	\par
	The contributions of this work are in summary: Our hybrid model-architecture and modified
	LIME-based method for generating explanations of classifications of audio voice samples as
	synthetic or bona-fide and the two metrics that we introduce in order to assess the impact of
	individual features over numerous local explanations.
	\par
	This paper is divided into the following sections:
	\begin{itemize}
		\item §\ref{sec:related} Related Work, where we discuss the most relevant related
			research in the field, in order to create a picture of the current state-of-the-art
		\item §\ref{sec:method} Method, where we describe the general model implementation, as
			well as the implementation of our LIME-based method for generating explanations in
			more detail.
		\item §\ref{sec:experimental_results} Experimental Results, where we discuss how we performed our
			experiments, how our hybrid model performs, as well as use our proposed metrics in
			order to assess the feature impact in the average explanation.
		\item §\ref{sec:discussion} Discussion, where we discuss the implication of the results of
			our experiments, including the potential of the explanations as well as the
			shortcomings and limitations of our method.
	\end{itemize}
	\section{Related Work}
	\label{sec:related}
	Several researchers have already followed similar or relevant directions to that of this work.
	This section will summarize some efforts in previous research divided into categories of
	perceptible and imperceptible features, as well as previous efforts at explainability in this
	field.
	\subsection{Synthetic Voice Detection With Imperceptible Features}
	The most common approach in this field makes use of either learned or hand-crafted
	imperceptible features. Examples include spectrographic features such as the mel-spectrogram
	and its hand-crafted derivative the mel-frequency cepstral coefficients (MFCCs), both of which
	are present in our method due to their widespread successful use.
	\par
	The work of Anagha et al. \cite{anagha_audio_2023} is an example that makes use of
	mel-spectrograms in combination with an architecture based on convolutional neural networks,
	achieving well-performing results on the ASVSpoof2019 dataset \cite{wang_asvspoof_2020}.
	\par
	\sloppy
	A further work from Yan et al. \cite{yan_initial_2022} makes use of MFCCs, in combination with
	other hand-crafted imperceptible features such as linear frequency cepstral coefficients
	(LFCCs) to not only classify audio samples as synthetic or authentic, but also to correctly
	classify the vocoder used to produce the sample with nearly perfect accuracy on a handcrafted
	dataset.
	\par
	Other researchers, such as in the work of Qais et al. \cite{qais_deepfake_2022} make use of
	Fourier transforms, such as the short term Fourier transform (STFT). The aforementioned
	example uses these features combined with others such as the aforementioned mel-spectrogram
	based features. They also pair these features with a convolutional architecture and achieve
	notably accurate results on the ASVSpoof2017 dataset \cite{delgado_asvspoof_2018}.
	\par
	Yang et al. \cite{yang_robust_2024} demonstrate a comparison of multiple features, as well as a
	feature selection method, with the goal of maximizing model efficiency. They run experiments
	with several hand-crafted features as well as learned features. They train and test using
	three datasets, ASVSpoof2019 \cite{wang_asvspoof_2020}, ASV2021 \cite{liu_asvspoof_2023} and
	In-the-Wild \cite{muller_does_2022}, presenting their results for each feature, as well as
	the results of their selection and classification fusion methods.
	\subsection{Synthetic Voice Detection With Perceptible Features}
	There is also a smaller but still notable body of work encompassing the use of perceptible
	features for the detection of synthetic voices, making use of a variety of features, some of
	which were selected for use in our experiments.
	\par
	The work of Barrington et al. \cite{barrington_single_2023} explores the idea of
	classification using perceptible features, highlighting their potential for improving
	explainability in the space of deepfake audio detection, although they only implement a
	classifier and not an explainer. They do, however, also demonstrate a drop in performance when
	using perceptible features in comparison with imperceptible hand-crafted and deep-learning
	features in their experiments, which were performed on a combination of synthetic and real
	audio datasets.
	\par
	Chaiwongyen et al. \cite{chaiwongyen_contribution_2022,chaiwongyen_deepfake-speech_2023} also
	approach using perceptible features for classification in their works. They developed a
	perceptron architecture with a single hidden layer and trained and tested using the dataset
	of the ADD2022 Challenge \cite{yi_add_2024}, resulting in lackluster performance in 2022, with
	better results using an expanded and improved feature set in 2023.
	\par
	Li et al. \cite{li_comparative_2022} also approach perceptible features in their work,
	combining them with imperceptible (referred to as physical in the work) features and using
	various neural networks in their experiments. With their combination of perceptible and
	imperceptible features, they were able to produce the best performance results of their
	experiments in comparison to experiments run with only perceptible or imperceptible features,
	which were performed as well on the dataset of the ASV2022 Challenge \cite{yi_add_2024}.
	\subsection{Explainable Models}
	There have also been some efforts in the previous research to pair synthetic voice detection
	models with explainable methods.
	\par
	One previous example of an implementation of an explainable model for use in audio deepfake
	detection came from Ge et al. \cite{ge_explaining_2024}, who explain feature influence on
	models using the SHAP (SHapley Additive exPlanations) method. They apply this method using
	log-scaled power spectrograms as a feature, training and testing using the datasets from the
	ASV2019 challenge \cite{wang_asvspoof_2020}, they are able to determine and graphically
	represent areas of importance on the spectrogram, as well as globally summarize the
	SHAP-values.
	\par
	Another example is presented by Haq et al. \cite{haq_multimodal_2023}, who use the changes in
	emotional state as an input feature and represent "unlikely" changes on a graph so that it can
	be understood by an eventual end user. In this case, they achieved their results by combining
	the output of fake video and fake audio classifiers in order to produce a final classification
	for a video sample with audio. Testing against the presidential deepfake dataset
	\cite{sankaranarayanan_presidential_nodate}, they achieve impressive results in comparison to
	the standing benchmark on the dataset at the time.
	\section{Method}
	\label{sec:method}
		\subsection{Black-Box Model}
		We created, trained and evaluated the model used for our experiments using the tools
		available in the TensorFlow and Keras libraries \cite{tensorflow2015-whitepaper}. This
		section will go into further detail about the experimental setup for the black-box model
		as well as the way in which explanations were created and summarized.
			\subsubsection{Features}
			In order to increase the likelihood of the explanations being useful and
			understandable to the end user, we placed a focus on using multiple perceptible
			features as inputs for the classifier. We added used two imperceptible features with
			the hypothesis, based on previous research \cite{barrington_single_2023,
			chaiwongyen_contribution_2022,chaiwongyen_deepfake-speech_2023,li_comparative_2022},
			that they would positively of increasing the model performance.
			\par
			Several features were used for the final version of the black-box model. Some are
			features widely used in research involving the detection of synthetic while others are
			more specific and are cited accordingly. The features that are not summarized for a
			whole sample are extracted using a sliding window method, preventing a loss of
			fidelity that can be caused by compression of the feature to a standard size. The
			features we considered perceptible are as follows:
			\begin{itemize}
				\item
					\sloppy
					Harmonic-to-noise ratios: A perceptible feature inspired by previous research
					\cite{chaiwongyen_contribution_2022,chaiwongyen_deepfake-speech_2023,
					li_comparative_2022} but done in a sliding window fashion instead of on the
					whole file, the HNRs are the ratios of the strengths of harmonic frequencies
					to the strengths of "noise", the total strength outside the harmonic
					frequencies. Unlike some previously seen examples
					\cite{chaiwongyen_contribution_2022, chaiwongyen_deepfake-speech_2023}, I
					calculate this ratio for each fundamental frequency length, instead of using
					one value for the whole sample. Given that \(\gamma_{i}\) is the harmonic
					energy in a given fundamental frequency cycle and \(\iota_{i}\) is the
					residual energy in a given fundamental frequency cycle, the HNR at that cycle
					\(hnr_{i}\) is calculated as follows:
					\[ hnr_{i} = 20log\frac{\gamma_{i}}{\iota_{i}}. \]
				\item
					Fundamental frequency lengths: Considered a perceptible feature, the
					fundamental frequency lengths (f0 lengths) are the lengths of every
					fundamental frequency cycle in the sample. This has also been previously used
					in certain research on synthetic audio detection \cite{xue_audio_2022}. The
					output is one-dimensional with time as the axis.
				\item
					Onset strength: As used in previous research \cite{li_comparative_2022}, this
					perceptible feature represents the strengths of each onset in the audio
					sample, where an onset is point where there is a sudden rise in energy across
					the audio spectrum. This results in a one-dimensional output with time as the
					axis.
				\item
					Intensity: Also inspired by previous research \cite{li_comparative_2022},
					intensity, also a perceptible feature, is the total power at each point in the
					audio sample, given in db. This is calculated by creating a fourrier
					transformation of the sample and then summing across the frequency-axis for
					every point on the time-axis. Resulting in a one-dimensional output.
				\item
					Pitch-fluctuations: Also classified as a perceptible feature, the pitch
					fluctuations are calculated for every sample in the audio as the difference
					between pitch at the given point and the pitch at the previous point. Pitch is
					estimated based on the maximum power harmonics. This feature is similar to the
					use of summarized pitch fluctuations in previous research
					\cite{khanjani_learning_2023}. This feature is also one-dimensional with time
					as the axis. Given that \(H_{i}\) is the set of harmonic frequencies at
					fundamental frequency cycle \(i\) with \(h_{i} \in H_{i}\) as a frequency of
					the set and \(s(h_{i})\) is the power of a given harmonic frequency at cycle
					\(i\), the pitch can be estimated as follows:
					\[p_{i} = s(max(H_{i})).\]
					Then, given an offset \(x\), the pitch fluctuation at cycle \(i\), \(pf_{i}\),
					can be calculated as follows:
					\[pf_{i} = p_{i}-p_{i-x}.\]
				\item
					Jitter features: As defined in previous attempts to identify synthetic audio
					\cite{chaiwongyen_deepfake-speech_2023} with perceptible features,
					jitter-based features measures the absolute variations in fundamental
					frequency cycles in comparison to the nearest x neighbors. The jitter of a
					sample in relation to the nearest \(x\) samples can be described as follows:
					\[ jitter(x) = \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|T_{i}
						(\frac{1}{x}\sum_{n=i-m}^{i+m}T_{n})|}
					{\frac{1}{N}\sum_{i=1}^{N}T_{i}}, \]
					where \(T_{i}\) represents the extracted fundamental frequency length at cycle
					\(i\), \(N\) is the number of fundamental frequency periods and \(m\) is
					\(\lfloor \frac{x}{2} \rfloor\)
				\item
					Shimmer features: Similarly defined to the jitter features, shimmer features,
					also considered perceptible and are used in the same research. They instead
					make use of the amplitudes at each fundamental frequency period. Their purpose
					is to capture irregular vocal fold vibrations which may be an indication but
					not a guarantee of a synthetic voice. The shimmer of a sample in comparison to
					the nearest \(x\) fundamental frequency cycles is calculated as follows:
					\cite{chaiwongyen_deepfake-speech_2023}.
					\[ shimmer(x) = \frac{ \frac{1}{N-1}\sum_{i=1}^{N-1}|A_{i}
						(\frac{1}{x}\sum_{n=i-m}^{i+m}A_{n})|}
					{\frac{1}{N}\sum_{i=1}^{N}A_{i}}, \]
					where \(A_{i}\) is the amplitude at cycle \(i\), \(N\) is the number of
					fundamental frequency cycles in the sample and \(m\) is once again 
					\(\lfloor \frac{x}{2} \rfloor\)
			\end{itemize}
			We also chose to include two perceptible features in our hybrid model, due to the
			performance-enhancing impact that perceptible features can have
			\cite{chaiwongyen_deepfake-speech_2023} when paired with perceptible features. These
			features are the mel-spectrogram and their derivative mel-frequency cepstral
			coefficients. We choose these features because they are well established in the
			research and have demonstrated consistent good performance in other examples,
			\cite{qais_deepfake_2022,anagha_audio_2023,fathan_mel-spectrogram_2022,
			altalahin_unmasking_2023,hamza_deepfake_2022,yan_initial_2022}. They allow for a
			spectrographic representation and analysis of an audio sample transformed in a way to
			better represent human perception of frequencies \cite{qais_deepfake_2022}.
			\subsubsection{Model Architecture}
			We designed our model architecture to allow the combination of diverse features into
			one final prediction, regardless of whether the individual features have the same
			shape. We achieved this by using a separate model for each feature, with its own input
			and output layers (further referred to as sub-models), which are then concatenated
			together and processed though a model that pools into a final output value (which we
			will further refer to as the terminus model). The modular structure allows not only
			better performance and less memory use, due to the fact that fewer transformations are
			required for the input preparation, but also allows for more flexibility in the
			construction of the model, making it easy to add and remove features using pre-defined
			functions depending on feature type. This kind of architecture also provides an
			advantage to the generation of explanations, as only the terminus part of the hybrid
			model must be explained to interpret the importance of each input feature. Further
			clarification is in §\ref{sec:method_exp}.
			\par
			Figure \ref{fig:model_plot} depicts the meta-structure of our hybrid model, with each
			feature first being processed in its own separate sub-model before being concatenated
			into the terminus model at the end, which produces the final result. Figure
			\ref{fig:submodel} offers a deeper look into the general architecture of a singular
			sub-model in our method. The sub-models begin with an input layer, taking either a
			vector or matrix depending on the dimensionality of the feature to which the model is
			tailored. The inputs are then passed into alternating convolutional and pooling
			layers, in order to increase the localized pattern detection capability within the
			sub-models. Figure \ref{fig:term_model} is a generalized representation of the
			terminus model, which has a classic perceptron architecture, with a varying number of
			hidden layers, depending on the experiment.
			\begin{figure}[htbp]
				\begin{center}
					\includegraphics[width=0.7\textwidth]{images/model_fig.png}
					\caption{The general architecture for the models used in these experiments,
					containing \(n\) features. The features are processed in their tailored
					sub-models before being processed together in the terminus model.}
					\label{fig:model_plot}
				\end{center}
			\end{figure}
			\begin{figure}[htbp]
				\begin{center}
					\includegraphics[width=0.7\textwidth]{images/submodel.png}
					\caption{A generalized representation of a sub-model, with the input as either
					a vector or a matrix followed by alternating convolutional and pooling layers,
					with the output as a flattened vector. Here, \(n\) is the number of
					convolutional-pooling pairs.}
					\label{fig:submodel}
				\end{center}
			\end{figure}
			\begin{figure}[htbp]
				\begin{center}
					\includegraphics[width=0.7\textwidth]{images/term_model.png}
					\caption{A generalized representation of the terminus model paired with \(n\)
					sub-models and containing \(m\) hidden layers.}
					\label{fig:term_model}
				\end{center}
			\end{figure}
			\subsubsection{Training and Evaluation}
			We trained the black-box model described in the previous section on the first portion
			of the dataset, with the exact size of the training set differing depending on the
			exact experiment (more details in §\ref{sec:experimental_results}). Evaluation was
			then performed on the remaining samples in the dataset. Training batches consisted of
			100 samples each with every window position for each sample and an upper limit of
			1,000,000 lines per batch. For each batch, we performed either one or two epochs.
			Because a single sample can have multiple input lines, evaluation cannot be performed
			on a per-line basis, but must instead be summarized. Two methods are possible: median
			and mean, with the difference in results being negligible in our experiments. The
			default threshold is 0.5, meaning that a result under 0.5 is a negative result and
			over 0.5 is a positive result. This threshold can however be adjusted, useful for
			computing certain metrics.
		\subsection{Explanations}
		\label{sec:method_exp}
		We generate the explanations using the previously described surrogate model with the Local
		Interpretable Model-agnostic Explanation (LIME) method \cite{ribeiro_why_2016}. This
		section will describe in detail how individual explanations are generated at a local level
		for individual samples, as well as how we evaluate the usefulness of the explanations in
		at a global scale.
			\subsubsection{Generation of the Explanations}
			Because of the multi-shaped nature and independence of the input layers of the model,
			we cannot generate the explanations directly from the input itself. However, because
			the sub-models are separate from one another, having no influence on each other before
			being concatenated at the terminus, only the terminus part of the model is relevant
			for assessing the importance of the features in a single evaluation. This does,
			however, present two challenges:
			\begin{itemize}
				\item There are multiple input rows per sample.
				\item The features cannot be used directly for assessment at the terminus input
					layer.
			\end{itemize}
			We solved the first problem by taking the mean of the explanation weights of each
			feature produced by each row of the sample features. This results in an average
			contribution of each feature to the classification of the model. Because the
			end-classification of the surrogate is also a mean, the means of the LIME-results are
			an accurate representation of the aggregate influence.
			\par
			We approached the second problem by generating intermediate data for the
			LIME-evaluation, the prediction outputs of the sub-models, in order to determine what
			the input of the terminus model would be. We generate intermediate data by first
			decomposing the surrogate model into its component models, the sub-models and the
			terminus, and for each sub-model, running a prediction on the given input features, so
			that we have the intermediate results of each sub-model before the processing of the
			terminus model. We stored the results of the predictions of the sub-models and used
			for further analysis. This was done not only for the sample explanation data but also
			for a random sampling of the training data, so that the LIME explainer could use this
			data to generate localized estimated of model behavior.
			\par
			We then summarized The results of the LIME explainer together on a per-feature basis
			and normalized them to produce a decimal number between -1 and 1 for every feature. In
			this case, a negative value implies that the given feature pushed the result of the
			model in the negative direction (i.e. not fake), while a positive value indicates the
			opposite.
			\subsubsection{Evaluation of the Explanations}
			In order to assess if the explanations have the potential to be useful to an end
			user, we pursued the goal of summarizing many explanations generated using the
			previously described method together in a meaningful way in order to make a conclusion
			about the general usefulness of the individual features and the influence and
			potential use of the perceptible features. However, due to a lack of existing
			standardized metrics for the aggregate, global-scale evaluation of local explanations,
			we propose two metrics, with which we can better assess the usefulness of our
			method.
			\par
			In order to contextualize the metrics, we will make the following definitions.
			\begin{itemize}
				\item Let \(S\) be the set of samples with \(s \in S\) as an element of the set.
				\item Let \(F\) be the set of features with \(f \in F\) as an element of the set.
				\item Let \(w_{fs}\) be the weight value of feature \(f\) of sample \(s\) as
					produced by the explainer.
				\item Let \(l_{s}\) be the correct label of the sample \(s\).
			\end{itemize}
			The first metric we will define is the mean of the absolute values of the weights of
			the explanations, summarized along the sample axis, delivering one value per feature.
			We will further refer to this metric as importance, and it can be mathematically
			defined as follows for every \(f \in F\):
			\[I(f) = \frac{\sum_{s \in S} |w_{fs}|}{|S|}.\]
			The second metric we will define is the average aggregate correctness on a per-feature
			basis, which we will further refer to as trust. The trust per feature \(f \in F\) can
			be defined as follows:
			\[T(f) = \frac{\sum_{s \in S} w_{fs}(2l_{s}-1)}{|S|}.\]
	\section{Experimental Results}
	\label{sec:experimental_results}
	Using the methods described in the previous section, we performed experiments training and
	testing the surrogate model as well as generating many explanations for the purpose of
	aggregate evaluation. In this section, we will first discuss the dataset that we used for our
	experiments, the results and performance of the surrogate model, followed by an evaluation of
	the aggregated explanations.
		\subsection{Dataset}
		The dataset used for these experiments is the In-the-Wild dataset, one which focuses on
		the generalization of audio deepfake detection models by collecting real-world data, in
		comparison to other examples that use more controlled laboratory conditions
		\cite{muller_does_2022}. We have chosen it for these experiments because of the
		aforementioned focus on generalization, its relative recency compared to some others and
		the fact that it has also been used previously by some other experiments, which provides a
		useful perspective against which we can compare our method.
		\subsection{Black-Box Model Results}
		\begin{table}[htbp]
			\caption{Summary of the evaluation results of the experiments}
			\vspace{10pt}
			\centering\notsotiny
			\begin{tabular}{c | c | c | c | c | c}
				\hline
				Terminus & Features & Training & Accuracy & EER & AUC \\
				\hline
				Simple & standard & 3474 & 95.02\% & 0.03702 & 0.90297 \\
				3 hidden layers & standard & 10000 & \textbf{96.27\%} & \textbf{0.03214} & 0.81763 \\
				Simple & standard & 23833 2 epochs & 90.07\% & 0.90069 & 0.78889 \\
				Simple & standard & 10000 2 epochs & 94.43\% & 0.04408 & 0.89445 \\
				Simple & standard & 10000 & 94.28\% & 0.04840 & 0.90182 \\
				3 convolutional layers & standard & 10000 & 62.80\% & 0.37198 & 0.50051 \\
				Simple & expand pitch\-flucs & 10000 & 93.31\% & 0.05661 & \textbf{0.92727} \\
				3 hidden layers & expand pitch\-flucs & 10000 & 93.87\% & 0.05317 & 0.83263
			\end{tabular}
			\label{table:eval-results}
		\end{table}
		\begin{figure}[htbp]
			\centering
			\includegraphics[width=0.7\textwidth]{images/roc_cterm.png}
			\caption{ROC curve of the best-performing model, with 3 hidden layers in the terminus
			and that standard feature set.}
			\label{fig:roc_cterm}
		\end{figure}
		\begin{figure}[htbp]
			\centering
			\includegraphics[width=0.7\textwidth]{images/roc_mpf.png}
			\caption{ROC curve of the model with expanded pitch fluctuation features.}
			\label{fig:roc_mpf}
		\end{figure}
		\begin{figure}[htbp]
			\centering
			\includegraphics[width=0.7\textwidth]{images/roc_mpf_cterm.png}
			\caption{ROC curve of the model with expanded pitch fluctuation features and a
			terminus with three hidden layers.}
			\label{fig:roc_mpf_cterm}
		\end{figure}
		\sloppy
		We provide summary of some experiments based on the method described above in table
		\ref{table:eval-results}. The table details the following information: The type of
		terminus model used, where a "simple" terminus model does not include hidden layers; the
		set of features used, where "standard" is the feature set as described above including
		only one pitch-fluctuation feature, with expanded pitch fluctuations being the same
		feature set with the addition of more pitch fluctuation features with different comparison
		distances; the training batch size and the number of epochs per batch if there were more
		than one; followed by several performance metrics. Figures \ref{fig:roc_cterm},
		\ref{fig:roc_mpf} and \ref{fig:roc_mpf_cterm} depict the ROC curve of three of the models,
		which were selected for use in explanation generation in the following section.
		\par
		The model that resulted in the best accuracy and EER had the standard feature set and was
		paired with a terminus with three hidden perceptron layers. We trained this model on the
		first 10000 samples of the dataset, and we evaluated it on the rest of the dataset using
		the methods previously described. It demonstrated an evaluation accuracy of 96.3\% with
		the default threshold of 0.5, as well as an EER of approximately 0.03214. The best AUC
		score, however, was achieved in another experiment, where the model had expended pitch
		fluctuations and a terminus model without hidden layers, with a value of 0.92727.
		\subsubsection{Comparison to Other Methods}
			\begin{table}
				\caption{A summary of the experimental results of other methods that have been
				evaluated against the same dataset. (When multiple experiments with one method
				were done, the best result was used for this table. The results only include
				evaluations on the full length samples, and not cropped versions, as we tested on
				the full-length samples)}
				\vspace{10pt}
				\centering
				\begin{tabular}{c | c}
					\hline
					Model Architecture & EER \\
					\hline
					STATNet \cite{ranjan_statnet_2022} & \textbf{0.00199} \\
					Fusion \cite{yang_robust_2024} & 0.2427 \\
					ASSERT \cite{yi_audio_2023} & 0.2473 \\
					Selection \cite{yang_robust_2024} & 0.2598 \\
					ResNet18 \cite{yang_robust_2024} & 0.2748 \\
					LCNN \cite{yi_audio_2023} & 0.3514 \\
					RawGAT-ST \cite{muller_does_2022} & 0.37154 \\
					MesoInception \cite{muller_does_2022} & 0.37414 \\
					GMM \cite{yi_audio_2023} & 0.3749 \\
					RawNet2 \cite{muller_does_2022} & 0.37819 \\
					Transformer \cite{muller_does_2022} & 0.43775 \\
					CRNNSpoof \cite{muller_does_2022} & 0.44500 \\
					RawPC \cite{muller_does_2022} & 0.45715 \\
					MesoNet \cite{muller_does_2022} & 0.46939 \\
					ResNet18 \cite{muller_does_2022} & 0.49759 \\
					LSTM \cite{muller_does_2022} & 0.53711 \\
					LCNN-LSTM \cite{muller_does_2022} & 0.61500 \\
					LCNN \cite{muller_does_2022} & 0.65559 \\
					LCNN-Attention \cite{muller_does_2022} & 0.66684 \\
				\end{tabular}
				\label{table:other_results}
			\end{table}
			In order to contextualize our results, we compare them to other results produced using
			tests on the same dataset. Table \ref{table:other_results} presents some experimental
			results that other methods have achieved on this dataset, sorted by source and model
			architecture. The results are given in EER, and the best result is in bold. In
			comparison to the other methods, our method performs very well, only beaten out by one
			other \cite{ranjan_statnet_2022}, which uses an expanded RawNet2-based network, taking
			the raw waveform as an input. They trained their method on 70\% of the dataset, used
			10\% for validation and the remaining 20\% for testing. This indicates a potential for
			raw-waveform based learned feature in the accurate detection of audio deepfakes, but
			does not offer a clear path for explainability, as the model does not make any
			intrinsic distinction of audio characteristics that humans are able to perceive. Our
			model remains, however, competitive, indicating that it can provide a good base for
			trustworthy explanations.
		\subsection{Explanation Results}
		\begin{table}[htbp]
			\caption{Summary of the aggregated explanation results of the best performing model}
			\vspace{10pt}
			\centering
			\begin{tabular}{c | c | c}
				\hline
				Feature & importance & trust \\
				\hline
				HNRs & 0.1140 & -0.0543 \\
				mel spectrogram & 0.1084 & -0.0378 \\
				MFCC & \textbf{0.7029} & \textbf{0.3753} \\
				f0 lengths & 0.0 & 0.0 \\
				onset strengths & 0.0 & 0.0 \\
				intensities & 0.0 & 0.0 \\
				pitch fluctuations & 0.0 & 0.0 \\
				jitter features & 0.1277 & 0.0250 \\
				shimmer features & 0.1227 & 0.0017
			\end{tabular}
			\label{table:exp-results-cterm}
		\end{table}
		\begin{figure}
			\centering
			\includegraphics[width=0.7\textwidth]{images/exp_cterm.png}
			\caption{Graphical representation of the aggregated explanation results of the best
			performing model}
			\label{fig:exp_cterm}
		\end{figure}
		\begin{table}[htbp]
			\caption{Summary of the aggregated explanation results of the model with expanded
			pitch fluctuation features}
			\vspace{10pt}
			\centering
			\begin{tabular}{c | c | c}
				\hline
				Feature & importance & trust \\
				\hline
				HNRs & 0.1072 & -0.0148 \\
				mel spectrogram & 0.0428 & -0.0187 \\
				MFCC & \textbf{0.4383} & \textbf{0.2345} \\
				f0 lengths & 0.0340 & -0.0057 \\
				onset strengths & 0.0101 & -0.0027 \\
				intensities & 0.0021 & 0.0021 \\
				pitch fluctuations & 0.0059 & 0.0020 \\
				jitter features & 0.2439 & 0.0019 \\
				shimmer features & 0.1664 & 0.0070
			\end{tabular}
			\label{table:exp-results-more-pitch-flucs}
		\end{table}
		\begin{figure}
			\centering
			\includegraphics[width=0.7\textwidth]{images/exp_mpf.png}
			\caption{Graphical representation of the aggregated explanation results of model with
			expanded pitch fluctuation features.}
			\label{fig:exp_mpf}
		\end{figure}
		\begin{table}[htbp]
			\caption{Summary of the aggregated explanation results of the model with expanded
			pitch fluctuation features and a terminus with three hidden layers.}
			\vspace{10pt}
			\centering
			\begin{tabular}{c | c | c}
				\hline
				Feature & importance & trust \\
				\hline
				HNRs & 0.2667 & -0.0788 \\
				mel spectrogram & 0.0982 & -0.0434 \\
				MFCC & \textbf{0.2879} & \textbf{0.2408} \\
				f0 lengths & 0.1690 & -0.0521 \\
				onset strengths & 0.1384 & -0.0613 \\
				intensities & 0.1576 & -0.0504 \\
				pitch fluctuations & 0.0 & 0.0 \\
				jitter features & 0.1444 & 0.0118 \\
				shimmer features & 0.1747 & 0.0116
			\end{tabular}
			\label{table:exp-results-mpf-cterm}
		\end{table}
		\begin{figure}
			\centering
			\includegraphics[width=0.7\textwidth]{images/exp_mpf_cterm.png}
			\caption{Graphical representation of the aggregated explanation results of the model
			with expanded pitch fluctuation features and a terminus with three hidden layers.}
			\label{fig:exp_mpf_cterm}
		\end{figure}
		In order to obtain a full picture of the quality of the explanations generated by this
		method, we have used selected three models from the last section and generated explanations
		on the first 500 samples of the testing data. The models selected were the model with the
		best accuracy and EER results, with the standard features and with three hidden layers in
		the terminus model, the model with expanded pitch-fluctuation features and a simple
		terminus, which achieved the best AUC result, as well as the model with expanded
		pitch-fluctuation features and a terminus with three hidden layers. We selected these
		models were selected because we wanted to evaluate the best performing models, as well as
		see how the changed pitch-fluctuation feature affected the evaluation of the explanations.
		We then aggregated and summarized the explanations generated using both of the previously
		introduced metrics. A summary of the results of both metrics summarized by the feature is
		presented in tables \ref{table:exp-results-cterm},
		\ref{table:exp-results-more-pitch-flucs} and \ref{table:exp-results-mpf-cterm}. The same
		statistics are also represented graphically in the figures \ref{fig:exp_cterm},
		\ref{fig:exp_mpf} and \ref{fig:exp_mpf_cterm} respectively.
		\subsubsection{Model with Standard Feature Set and Complex Terminus}
		Measuring by both metrics, the MFCCs had the most positive and the most correct influence
		on the classification results of the model, With the jitter features as a distant second
		for trust and importance. This indicates that the MFCCs had not only the most influence,
		but also the most correct influence on the result of the classification, while the other
		features had either little, no or negative influence on the classification. Jitter
		features are, however, still notable for being perceptible features with a positive trust
		value.
		\subsubsection{Model with Expanded Pitch-Fluctuation Features}
		This model had comparable results, with the main difference being that all metrics have
		non-zero values, meaning that all the present features had an influence on the
		classifications which were used to generate the explanations. As with the previous model,
		HNRs and mel-spectrograms had negative overall trust values, with the addition of the
		onset strengths also presenting a negative value, implying that these features have caused
		more harm than good in the tested classifications. In this example, both jitter and
		shimmer features are notable for their positive trust values indicating their potential as
		trustworthy perceptible features for explanations.
		\subsubsection{Model with Expanded Pitch-Fluctuation Features and Complex Terminus}
		The model with multiple pitch-fluctuation features in combination with a terminus
		containing multiple hidden layers fared similarly to the other two, this time with the
		pitch fluctuations having no influence on the outcome whatsoever. Unlike the previous
		examples, the trust value of the intensities feature was negative here. The trust of the
		jitter and shimmer features, however, remain positive.
	\section{Discussion}
	\label{sec:discussion}
	In this section, we will discuss the implications of the experimental results of the previous
	sections with a special focus being placed on the results of the explanation tests as it
	relates to the potential explainability for potential end users of such a system.
		\subsection{Viability and Potential of the Explanations}
		Because of these models' inclusion of several perceptible features, we made the hypothesis
		that this method has the potential to produce explanations that can be used as the basis
		for results more presentable to an end user. As can be seen in the previous section in
		the tables \ref{table:exp-results-cterm} \ref{table:exp-results-more-pitch-flucs}, the
		perceptible features did not contribute as much to the end result of the classification as
		their imperceptible counterparts, in spite of their increased presence in the overall set
		of input features. Two features, the HNRs and the mel-spectrograms even consistently had a
		net-negative impact on the correctness of the classifications in the experiments,
		according to our trust metric, leaving the question open, if future experiments may
		perform better without these features. In the case of the model without expanded
		pitch-fluctuations, four perceptible features, f0 lengths, onset strengths, intensities
		and pitch fluctuations had an overall average influence on the classification of 0,
		measured on both metrics, indicating that the model did not learn a meaningful correlation
		between these features and the authenticity of audio samples in the dataset. These
		features did have an impact on the performance of the model with expanded pitch
		fluctuations, but the difference was still minimal. In the case of perceptible features
		with a positive trust score, the potential remains for their potential use in
		understandable explanations as, even though the classification needle of the model was not
		moved significantly by these features, they still demonstrate a certain reliability. In
		spite of this, we believe that it is still worth investigating if other kinds of
		perceptible features will have more of an influence on the overall classification or if
		there are other imperceptible features that could be used in the place of MFCCs that may
		provide a better balance between classification correctness and a proportional influence
		on the end result, so that the usefulness of the explanations can be even stronger than
		the method using the above feature combinations, and we believe that this kind of model
		architecture paired with LIME and evaluated with the presented metrics has the potential
		to be a backbone for future research.
		\subsection{Limitations}
		In spite of the potential that we see with this method, it does also present some
		limitations when it comes to its real-world use, aside from the shortfalls of the
		perceptible features discussed in the previous section. The first such limitation is in
		the computational performance of the model. Because of its relative size, the model
		requires relatively capable hardware in addition to several gigabytes of memory in order
		to classify samples. This means that the devices this or a similar model can run on are
		limited, and the possibility of running this or a similar model locally on a handheld
		device such as a smartphone is limited to none at the time of writing. Another limitation
		with this method is that, even though it is known that perceptible features such as the
		ones used in our method are able to be heard by the human ear and are even used in
		medicine, for example, for diagnoses \cite{chaiwongyen_deepfake-speech_2023}, it is not
		yet clear to what extent exactly these features can be understood by the average person.
		Cursory research \cite{warren_better_2024,sharevski_blind_2024} is already present into
		what factors humans use to identify fake and real audio samples, as well as their
		performance, but the named aspects of the samples have not yet been mapped to such
		perceptible features as the ones used in this method, leaving the question of the
		usability of such explanations open, in addition to the question of what additional
		perceptible features can possibly be extracted to mirror what people have specifically
		identified in fake voice samples. Finally, we have not addressed in our experiments how
		the features can be presented to the end user. Certain implementations of LIME already
		have means to present their results graphically, but the effectiveness of such methods has
		yet to be studied.
	\section{Conclusion}
	To conclude, we have presented a model architecture using a combination of various features,
	perceptible and imperceptible, with the hypothesis that such a model could be used to generate
	explanations of its own results using the LIME method that are more useful to a potential end
	user who is not already familiar with the field. This model architecture resulted in a high
	level of performance in comparison to several other methods, indicating the potential for
	consistent and correct classifications. The explanations, however, placed most of the weight
	on the selected imperceptible features, with most perceptible features either having little,
	no or negative impact on the end classification with the exception of jitter and shimmer,
	which offer potential to be used in explanations.
	\par
	We believe that this is a method that offers a large potential, but is not yet in a fully
	mature state. We suggest, however, that this method could be changed and improved in the
	future with other features or other changes to base parameters that could potentially produce
	more useful explanations while either holding the accuracy high like it is now or even
	improving it further.
	\newpage
	\sloppy
	\printbibliography
\end{document}